Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-06-14 parallel-tractability-J-v9/response2.tex
added some content in Response2

Hence, while the algorithms (e.g.\ Algorithm $\mathsf{A}_\text{bsc}$
or (or $\mathsf{A}_\text{opt}$) involve two inputs (rules and facts),
for the complexity analysis, we assume the rules part of the input to
be fixed. In general, however, the algorithms can handle different
rule sets. Please note that we still have to show taht the (data)
complexity holds for all possible sets of facts.

@@

\textcolor{red}{Why do you write $e(v_1, v_2)$ for edges and not just
  $(v_1, v_2)$?}

\textcolor{red}{In the materialization graph for Example~5, how is
  $A(a)$ derived from the ontology and $Q(a, b)$? }

@@

@@

1) Suppose that there is a one-to-one correspondence between
processors and rule instances, and each processor can access the
location of any atom in constant time (given as the parallel
assumption in Section~3.3).  \textcolor{red}{What might be confusing
  is the term rule instance. In Sec.~2.1, we do not introduce the
  term, but we say `ground instantiation of a rule`, which I suppose
  means the same thing, but maybe the reviewer reads `rule instance`
  as `rule', which is a significant difference. If `rule instance'
  indeed means `ground instantiation of a rule', we should unify our
  terminology to avoid confusion. } Since a processor $p$ corresponds
to a rule instance, e.g., $B_1, \ldots, B_n \to H$, processor $p$
knowns the relationship among $B_1, \ldots, B_n$ and $H$. Further,
processor $p$ knows the locations of the nodes $B_1, \ldots, B_n$ and
$H$. Thus, it could check the derivability of $B_1, \ldots, B_n$, and
bridge \textcolor{red}{`to bridge' means to connect in the sense of
  adding the edge to $G$?} $B_i$ and $H$ ($1 \leq i \leq n$) in
constant time instead of searching the whole graph $G$.

2) How to establish such a one-to-one correspondence between processors 
and rule instances and the constant-time access to each atom for processors?

We should also introduce the notion of Boolean circuit, which is a
circuit that consists of logic gates, e.g., AND and OR gates (see
\cite[Section~2.3]{Raymond95}).  A materialization graph can also be
seen as a Boolean circuit by treating the nodes as AND gates.\\
%
First, the evaluation of a Datalog program is equivalent to the evaluation 
of its corresponding Boolean circuit in terms of computational
complexity \cite{DMRT14a}. \\
%
Second, a Boolean circuit can be simulated by PRAM based on an
algorithm called parallel prefix computation \cite{LaFi80a}.  Further,
the complexity of simulation is lower than that of the evaluation of
the Boolean circuit \cite[Lemma~2.4.1]{Raymond95}.\\
%
Third, during the above simulation procedure, a one-to-one
correspondence between each gate in the Boolean circuit and the
processors in PRAM and the constant-time access are established.\\
%
In our case, the one-to-one correspondence between processors and rule
instances and the constant-time access to each node can also be
established by PRAM similarly to the above simulation procedure.

\textcolor{red}{We have added the above argumentation also to the
  paper. (needed?, TBD)}

>>>>

Hence, while the algorithms (e.g.\ Algorithm $\mathsf{A}_\text{bsc}$
or (or $\mathsf{A}_\text{opt}$) involve two inputs (rules and facts),
for the complexity analysis, we assume the rules part of the input to
be fixed. In general, however, the algorithms can handle different
rule sets. Please note that we still have to show that the (data)
complexity holds for all possible sets of facts.

@@

@@

In the submission, we used both of the two terms `rule instance'
and `ground instantiation of a rule', which denote the same thing.
We now unify the term by `rule instantiation' which is introduced
in Sec.~2.1. A rule instantiation is different from a rule.
A rule instantiation has no variable while a rule can have variables.
Checking the applicability of a rule depends on the size of the
processed graph. However, checking the applicability of a rule
instantiation only depends on the derivability of the ground atoms
in this rule instantiation. Suppose that a processor can access the
location of any atom in constant time. It can also check the applicability of
a rule instantiation in constant time.

In the last response, we gave another two kinds of problems, i.e.,
the computation of transitive closure and the matrix manipulation (
see \cite[Part~\uppercase\expandafter{\romannumeral2}~A]{Raymond95}. 
We made a mistake to use `matrix multiplication'
in the last response), to show that the constant time assumption
is also applicable. The first reference in the last response gives the related works
about this issue but not covering the details. It is also hard to find
the clues in the second reference.
In the following, we give a more detailed explanation.

@@

1) Suppose that there is a one-to-one correspondence between
processors and rule instantiations, and each processor can access the
location of any atom in constant time (given as the parallel
assumption in Section~3.3). Since a processor $p$ corresponds
to a rule instantiation, e.g., $B_1, \ldots, B_n \to H$, processor $p$
knowns the relationship among $B_1, \ldots, B_n$ and $H$. Further,
processor $p$ knows the locations of the nodes $B_1, \ldots, B_n$ and
$H$. Thus, it could check the derivability of $B_1, \ldots, B_n$, and
add an edge from $B_i$ to $H$ ($1 \leq i \leq n$) in
constant time instead of searching the whole graph $G$.

2) How to establish such a one-to-one correspondence between processors
and rule instantiations and the constant-time access to each atom for processors?

We should also introduce the notion of Boolean circuit, which is a
circuit that consists of logic gates, e.g., AND and OR gates (see
\cite[Section~2.3]{Raymond95}).  A materialization graph can also be
seen as a Boolean circuit by treating the nodes as AND gates.\\
%
First, the evaluation of a Datalog program is equivalent to the evaluation
of its corresponding Boolean circuit in terms of computational
complexity \cite{DMRT14a}. \\
%
Second, a Boolean circuit can be simulated by PRAM based on an
algorithm called parallel prefix computation \cite{LaFi80a}.  Further,
the complexity of simulation is lower than that of the evaluation of
the Boolean circuit \cite[Lemma~2.4.1]{Raymond95}.\\
%
Third, during the above simulation procedure, a one-to-one
correspondence between each gate in the Boolean circuit and the
processors in PRAM and the constant-time access are established.\\
%
In our case, the one-to-one correspondence between processors and rule
instantiations and the constant-time access to each node can also be
established by PRAM similarly to the above simulation procedure.

\textcolor{red}{We have added the above argumentation also to the
  paper. (needed?, TBD) \emph{Rephrase the relevant paragraphs in a brief way?}}
